Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-09-09 psy_peds.Rmd
optimizing hospitalisation sequences count

```{r}
# Adjusting patient age to the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
clinical_raw <- left_join(clinical_raw, demo_raw[c("patient_num", "cohort", "age")], by = c("patient_num", "cohort"))
clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)

## Adding columns with hospitalization number and hospitalization length to the clinical raw file
```{r}
count_sequences_hospitalisation <- function(df, ...) {
  # TODO: To be optimized
  total_span <- data.frame(total_span = seq(min(df$days_since_admission),
                                            max(df$days_since_admission))
  ) %>% 
    left_join(df[,c("in_hospital", "days_since_admission")],
              by = c("total_span" = "days_since_admission")
    ) %>% replace_na(list(in_hospital = 0))
  count_sequences <- rle(total_span$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  df[df$in_hospital == 1, "n_hospitalisation"] <- sequences
  df[df$in_hospital == 1, "len_hospitalisation"] <- sequences_len
  return(df)
}
clinical_raw <- clinical_raw %>%
  group_by(patient_num, cohort) %>% 
  group_modify(count_sequences_hospitalisation)
```

@@

diagnosis_table1 <- summary_diagnosis %>% 
    bind_rows(
    sex_by_period %>%
      group_by(period) %>% 
      dplyr::summarise(concept_code = "Total", count = sum(count))
  ) %>%
  pivot_wider( names_from = period, values_from = count) %>% 
  mutate(diag = paste(concept_code, description, sep = ": ")) %>% 
  select(diag, during_pandemic) %>% 
  rename(name = diag)

disorder_table1 <- summary_disorder_group_by_period %>% 
  bind_rows(
    dplyr::summarise(., disorder_group="Total", across(where(is.numeric), sum))
  ) %>% 
  mutate(result = format_counts(count, percentage)) %>% 
  select(-c(count, percentage)) %>% 
  pivot_wider( names_from = period, values_from = result ) %>% 
  # replace_na( list(early_pandemic = "0 (0 %)", 
  #                  late_pandemic = "0 (0 %)")) %>% 
    replace_na( list(during_pandemic = "0 (0 %)")) %>% 
  rename(name = disorder_group)

>>>>

```{r}
# Adjusting patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
clinical_raw <- left_join(clinical_raw, demo_raw[c("patient_num", "cohort", "age")], by = c("patient_num", "cohort"))
clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
```

## Adding columns with hospitalization number and hospitalization length to the clinical raw file
```{r}
count_sequences_hospitalisation <- function(df, ...) {
  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
                                                        max(df$days_since_admission))
  ) %>%
    left_join(df, by = c("total_span" = "days_since_admission")) %>%
    replace_na(list(in_hospital = 0))
  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  stopifnot(length(df$days_since_admission) == length(sequences))
  data.frame(days_since_admission = df$days_since_admission,
             n_hospitalisation = sequences,
             len_hospitalisation = sequences_len)
}
stopifnot(all(clinical_raw$in_hospital == 1))
hospitalisations_seq_df <- clinical_raw %>%
  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
  group_by(patient_num, cohort) %>%
  group_modify(count_sequences_hospitalisation)
clinical_raw <- left_join(clinical_raw,
                          hospitalisations_seq_df,
                          by = c("patient_num", "cohort", "days_since_admission"))
```

@@

diagnosis_table1 <- summary_diagnosis %>% 
    bind_rows(
    sex_by_period %>%
      group_by(period) %>% 
      dplyr::summarise(concept_code = "Total", count = sum(count))
  ) %>%
  pivot_wider( names_from = period, values_from = count) %>% 
  mutate(diag = paste(concept_code, description, sep = ": ")) %>% 
  select(diag, before_pandemic, during_pandemic) %>% 
  rename(name = diag)

disorder_table1 <- summary_disorder_group_by_period %>% 
  bind_rows(
    dplyr::summarise(., disorder_group="Total", across(where(is.numeric), sum))
  ) %>% 
  mutate(result = format_counts(count, percentage)) %>% 
  select(-c(count, percentage)) %>% 
  pivot_wider( names_from = period, values_from = result ) %>% 
  replace_na( list(during_pandemic = "0 (0 %)")) %>% 
  rename(name = disorder_group)
